Fluid-mechanical erosion creates anisotropic porous media

### 2022-06-02 docs/PhysicsPaper/StokesErosion.tex
NM

The embedded bodies may erode in response to the shear stresses induced by the intervening fluid flow. Erosion typically occurs over much longer timescales than does the fluid flow, permitting a {\em quasi-steady} approximation. In this approximation, the configuration of bodies is held fixed in order to compute the {\em steady} Stokes flow determined by \eqref{eqn:StokesEq}, and then this flow field determines the stresses acting to erode each body. We employ an erosion law in which the local rate of material loss is linearly proportional to the magnitude of the shear stress $\tau$ acting on the surface (CITE previous erosion work). The material loss gives rise to an inward velocity of the solid surface, $\Vn$, pointing in the direction normal to the surface. The erosion law is thus expressed as
\begin{align}
\Vn = \CE \, \abs{\tau}, 
	&\hspace{20pt} \xx \in \gamma, &&\mbox{\em erosion model}, \\
\tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
	&\hspace{20pt} \xx \in \gamma, &&\mbox{\em shear stress}.
  \label{eqn:shearStress}
\end{align}
where $\nn$ is a unit normal vector pointing into each body, $\ss$ is
a unit tangent vector pointing in the counterclockwise direction, and $\CE$ is a material-dependent erosion constant.

@@

The permeability of the medium is directly related to the total drag exerted by the collection of bodies. The stress tensor associated with the Stokes equations \eqref{eqn:StokesEq} is given by $\stress = -p \bvec{I} + \mu \left( \grad \uu + \grad \uu^T \right)$, and an alternate expression of \eqref{eqn:StokesEq} is $\grad \cdot \stress = 0$. Integrating this form of the equations over an arbitrary domain $D$ and applying the divergence theorem gives
\begin{equation}
\label{divthm}
0 = \int_{D} \grad \cdot \stress \, dV 
= \int_{\partial D} \stress \cdot \nn \, ds
\end{equation}
Consider $D \subseteq [-x_0, x_0] \times [-1, 1]$ to be the subset consisting of the fluid region (i.e.~excluding solid bodies), where $x_0 \ge 1$ will be chosen to include the entire porous region plus some amount of the buffer region. The boundary $\partial D$ consist of $\gamma$, the solid boundaries of all of the bodies, as well as an outer boundary. The total drag exerted by all of the solid bodies is given by
\begin{equation}
\FD = -\int_{\gamma} \stress \cdot \nn \, ds
\end{equation}

\nick{Do I want plus or minus? I decided I want the drag exerted BY the bodies and so this should point to the left.}

@@

Consider $\Omega$ to be inside $[-x_0, x_0] \times [-1, 1]$, where we will choose $x_0 > 1$ so that $\Omega$ includes the entire porous region.
The boundary integral on the right-hand-side of \eqref{divthm} has contributions from the boundaries of the erodable bodies $\gamma$ and from the outer boundary $\Gamma$, since $\partial \Omega = \gamma \cup \Gamma$. In particular, the total drag exerted by the erodable bodies is given by
\begin{equation}
\FD = -\int_{\gamma} \stress \cdot \nn \, ds
\end{equation}
which includes both pressure and viscous drag. The minus sign is due to the orientation of the normal vector. We will mostly be interested in the horizontal drag $\FD \cdot \ex$. \\

\nick{Need to check sign on drag and check whether I want the drag {\em on} the bodies or exerted {\em by} the bodies.} \\

>>>>

The embedded bodies may erode in response to the shear stresses induced by the intervening fluid flow. Erosion typically occurs over much longer timescales than does the fluid flow, permitting a {\em quasi-steady} approximation. In this approximation, the configuration of bodies is held fixed in order to compute the {\em steady} Stokes flow determined by \eqref{eqn:StokesEq}, and then this flow field determines the stresses acting to erode each body. We employ an erosion law in which the local rate of material loss is linearly proportional to the magnitude of the shear stress $\tau$ acting on the surface (CITE previous erosion work). The material loss gives rise to an inward velocity of the solid surface, $\Vn$, pointing in the direction normal to the surface. The erosion law is thus expressed as
\begin{align}
\Vn = \CE \, \abs{\tau}, 
	&\hspace{20pt} \xx \in \gamma, &&\mbox{\em erosion model}, \\
\tau = -\mu \left( ( \nabla \uu + \nabla \uu^T) \nn \right) \cdot \ss
	&\hspace{20pt} \xx \in \gamma, &&\mbox{\em shear stress}.
  \label{eqn:shearStress}
\end{align}
where $\nn$ is a unit normal vector pointing into each body, $\ss$ is
a unit tangent vector pointing in the counterclockwise direction, and $\CE$ is a material-dependent erosion constant.

@@

The permeability of the medium is directly related to the drag exerted by the collection of bodies. The stress tensor associated with the Stokes equations \eqref{eqn:StokesEq} is given by $\stress = -p \bvec{I} + \mu \left( \grad \uu + \grad \uu^T \right)$, and the Stokes equations can alternatively be expressed as $\grad \cdot \stress = 0$. Integrating over an arbitrary domain $D$ and applying the divergence theorem gives
\begin{equation}
\label{divthm}
0 = \int_{D} \grad \cdot \stress \, dV 
= \int_{\partial D} \stress \, \nn \, ds
\end{equation}
Consider $D \subseteq [-x_0, x_0] \times [-1, 1]$ to be the subset consisting of the fluid region (i.e.~excluding solid bodies), where $x_0 \ge 1$ will be chosen to include the entire porous region plus some amount of the buffer region. The boundary $\partial D$ consist of all solid-body boundaries $\gamma$, along with an outer boundary.

The hydrodynamic drag on the collection of bodies is obtained by integrating the surface traction over the boundary $\gamma$. The surface traction is given by 
\begin{equation}
- \stress \, \nn = p \nn + \tau \ss
\end{equation}
where the negative sign is a consequence of choosing the normal vector $\nn$ pointing {\em into} the bodies. The total drag on the collection of bodies is thus
\begin{equation}
\FD =  - \int_{\gamma} \stress \, \nn \, ds =  \int_{\gamma} p \nn + \tau \ss \, ds
\end{equation}

% NOTE ON SIGN OF DRAG
% The drag ON the bodies is F = \int p n + tau s. This is consistent with the JCP paper, with my code, and I have verified the calculation. This drag vector points to the right since it is the drag ON the bodies.

@@

The boundary integral on the right-hand-side of \eqref{divthm} has contributions from the boundaries of the erodable bodies $\gamma$ and from the outer boundary $\Gamma$, since $\partial \Omega = \gamma \cup \Gamma$. In particular, the total drag exerted by the erodable bodies is given by
\begin{equation}
\FD = -\int_{\gamma} \stress \cdot \nn \, ds
\end{equation}
We will mostly be interested in the horizontal drag $\FD \cdot \ex$. \\
